### 2022-06-09 Manuscript/src/main.tex
feat: cleaned code and started writing paper

%! Author = sbbfti
%! Date = 8/06/2022

% Preamble
\documentclass[11pt]{article}

\title{PMV model comparison}
\author{Federico Tartarini, Stefano Schiavon}

\usepackage{mypreamble}

\usepackage[printonlyused]{acronym}

\begin{document}

    \maketitle

    \begin{abstract}
        The \ac{pmv} model developed by Fanger predicts thermal sensation.
        It is used in its original formulation in the ISO 7730:2005, while the ASHRAE 55:2020 uses a modified version \acs{pmv-ce} to better account for air speeds.
        We compared their results against XXX thermal sensation votes collected in buildings.
    \end{abstract}

    \section*{Nomenclature}
    \input{myacronyms.tex}

    \section{Introduction}\label{sec:introduction}
    In 1970, Fanger~\cite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the ISO 7730:2005 Standard~\cite{iso7730} in its original form.
    The \ac{pmv} is the most extensively used thermal comfort index for estimating the thermal sensation of a group of people sharing the same thermal environment.
    It is used both by researchers and practitioners.
    However, Cheung et al. (2019) demonstrated that the accuracy of the \ac{pmv} in predicting self-reported thermal sensation of people is only \qty{34}{\percent}~\cite{Cheung2019}.

    ASHRAE-55 2020 Standard utilizes a modified version of the PMV model [4] that calculates a Cooling Effect (CE) using the Standard Effective Temperature (SET) equation for air speeds higher than \qty{0.1}{\m\per\s}.
    The value of CE is then subtracted from both the measured dry-bulb air temperature (tdb) and \ac{tr} and these results become the input into the PMV model.
    Since the CE already accounts for convective heat losses from the person to the environment, the air speed used to calculate the PMV value is set to 0.1 m/s.

    \printbibliography

\end{document}