### 2022-06-09 Manuscript/src/myacronyms.tex

%! Author = federico
%! Date = 10/06/2020

%\section*{Nomenclature}
\renewcommand{\baselinestretch}{0.75}\normalsize
\renewcommand{\aclabelfont}[1]{\textsc{\acsfont{#1}}}
\begin{acronym}[longest]

    \acro{tdb}[$t_{db}$]{dry-bulb air temperature\acroextra{, \unit{\celsius}}}
    \acro{rh}[$RH$]{relative humidity\acroextra{, \%}}
    \acro{v}[$V$]{average air speed\acroextra{, m/s}}
    \acro{tr}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, \unit{\celsius}}}
    \acro{clo}[$I_{cl}$]{total clothing insulation\acroextra{, clo}}
    \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}

    \acro{pmv}[PMV]{Predicted Mean Vote}
    \acro{pmv-ce}[PMV$_{CE}$]{PMV calculated in accordance with the ASHRAE 55:2020}
    \acro{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
    \acro{set}[SET]{Standard Effective Temperature\acroextra{, \unit{\celsius}}}
    \acro{ce}[CE]{Cooling Effect\acroextra{, \unit{\celsius}}}

\end{acronym}
\renewcommand{\baselinestretch}{1}\normalsize