### 2022-06-09 Manuscript/src/main.tex

%! Author = sbbfti
%! Date = 10/06/2020

\documentclass[review, 3p]{elsarticle}
%\documentclass[final]{elsarticle}
%\documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{mypreamble}

\begin{document}

    \input{sections/titleAbstract}

    \input{myacronyms.tex}

    \linenumbers

    \input{sections/introduction}

    \input{sections/methodology}

    \input{sections/results}

    \input{sections/discussion}

    \input{sections/conclusions}

    \bibliography{mybibfile}

\end{document}