### 2022-06-09 Manuscript/src/myacronyms.tex

%! Author = federico
%! Date = 10/06/2020

\section*{Nomenclature}
\renewcommand{\baselinestretch}{0.75}\normalsize
\renewcommand{\aclabelfont}[1]{\textsc{\acsfont{#1}}}
\begin{acronym}[longest]

    \acro{t-db}[$t_{db}$]{dry-bulb air temperature\acroextra{, $^{\circ}$C}}
    \acro{t-wb}[$t_{wb}$]{wet-bulb air temperature\acroextra{, $^{\circ}$C}}
    \acro{ti}[$t_{i}$]{indoor air temperature\acroextra{, $^{\circ}$C}}
    \acro{tout}[$t_{out}$]{outdoor air temperature\acroextra{, $^{\circ}$C}}
    \acro{t-op}[$t_{o}$]{operative air temperature\acroextra{, $^{\circ}$C}}
    \acro{t-cl}[$t_{cl}$]{clothing temperature\acroextra{, $^{\circ}$C}}
    \acro{tg}[$t_{g}$]{globe temperature\acroextra{, $^{\circ}$C}}
    \acro{rh}[$RH$]{relative humidity\acroextra{, \%}}
    \acro{v}[$V$]{average air speed\acroextra{, m/s}}
    \acro{t-r}[$\overline{t_{r}}$]{mean radiant temperature\acroextra{, $^{\circ}$C}}
    \acro{clo}[$I_{cl}$]{total clothing insulation\acroextra{, clo}}
    \acro{i-cl}[$i_{cl}$]{permeation efficiency of water vapor through the clothing layer}
    \acro{met}[$M$]{rate of metabolic heat production\acroextra{, W/m\textsuperscript{2}}}

    \acro{pmv}[PMV]{Predicted Mean Vote}
    \acro{ppd}[PPD]{Predicted Percentage of Dissatisfied\acroextra{, \%}}
    \acro{set}[SET]{Standard Effective Temperature\acroextra{, $^{\circ}$C}}
    \acro{ce}[CE]{Cooling Effect\acroextra{, $^{\circ}$C}}
    \acro{phs}[PHS]{Predicted Heat Strain}

    \acro{BMS}[BMS]{Building Management System}
    \acro{HVAC}[HVAC]{Heating, Ventilation, and Air Conditioning}
    \acro{VAV}[VAV]{Variable Air Volume}
    \acro{AHU}[AHU]{Air Handling Unit}

    \acro{p-sk}[$p_{sk,s}$]{water vapor pressure at skin\acroextra{, kPa}}
    \acro{p-a}[$p_{a}$]{water vapor pressure in ambient air\acroextra{, kPa}}

    \acro{wmo}[WMO]{World Meteorological Organization}
    \acro{who}[WHO]{World Health Organization}
    \acro{cdc}[CDC]{Centers for Disease Control and Prevention}
    \acro{noaa}[NOAA]{National Oceanic and Atmospheric Administration}
    \acro{epa}[EPA]{United States Environmental Protection Agency}
    \acro{iea}[IEA]{International Energy Agency}
    \acro{un}[UN]{United Nations}

\end{acronym}
\renewcommand{\baselinestretch}{1}\normalsize