### 2022-06-09 Manuscript/src/myglossary.tex
feat: added glossary

\usepackage[nonumberlist,nogroupskip]{glossaries}

\setglossarystyle{super}

\newglossaryentry{ipm}{
    name={Incremental PM},
    description={Incremental Personal Model}
}

\newglossaryentry{gpm}{
    name={Warm start PM},
    description={Warm start Personal Model}
}

\newglossaryentry{pm}{
    name={PM},
    description={Personal Model}
}
\newglossaryentry{ib}
{
name={iButton\textsuperscript{\tiny\textregistered}},
description={Wireless temperature sensors},
plural={iButtons\textsuperscript{\tiny\textregistered}},
firstplural={iButtons\textsuperscript{\tiny\textregistered}}
}

\newglossaryentry{ubibot}
{
name={UbiBot\textsuperscript{\tiny\textregistered}},
description={Wireless environmental sensor},
}

\newglossaryentry{7730}
{
name={ISO~7730:2005},
description={ISO~7730:2005 is a thermal comfort standard developed by ISO}
}

\newglossaryentry{55}
{
name={ASHRAE~55--2020},
description={ASHRAE~55--2020 is a thermal comfort standard developed by ANSI and ASHRAE}
}

\makenoidxglossaries