### 2022-06-09 Manuscript/src/myglossary.tex

\usepackage[nonumberlist,nogroupskip]{glossaries}

\setglossarystyle{super}

\newglossaryentry{7730}
{
name={ISO~7730:2005},
description={ISO~7730:2005 is a thermal comfort standard developed by ISO}
}

\newglossaryentry{55}
{
name={ASHRAE~55--2020},
description={ASHRAE~55--2020 is a thermal comfort standard developed by ANSI and ASHRAE}
}

\newglossaryentry{rhrn}
{
name={right-here-right-now},
description={Right here right now thermal comfort survey}
}

\newglossaryentry{pmv-ce}
{
name={PMV\textsubscript{CE}},
description={PMV calculated in accordance with the ASHRAE~55--2020}
}

\newglossaryentry{db2}
{
name={Comfort DB},
description={ASHRAE Global Thermal Comfort Database II}
}

\makenoidxglossaries