### 2022-06-20 Manuscript/src/myglossary.tex
paper: minor edits

\usepackage[nonumberlist,nogroupskip]{glossaries}

\setglossarystyle{super}

\newglossaryentry{7730}
{
name={ISO~7730:2005},
description={ISO~7730:2005 is a thermal comfort standard developed by ISO}
}

\newglossaryentry{55}
{
name={ASHRAE~55--2020},
description={ASHRAE~55--2020 is a thermal comfort standard developed by ANSI and ASHRAE}
}

\newglossaryentry{rhrn}
{
name={right-here-right-now},
description={Right here right now thermal comfort survey}
}

\newglossaryentry{pmv-ce}
{
name={PMV\textsubscript{CE}},
description={PMV calculated in accordance with the ASHRAE~55--2020}
}

\newglossaryentry{epmv}
{
name={ePMV},
description={Adjusted Predicted Mean Votes with Expectancy Factor}
}

\newglossaryentry{apmv}
{
name={aPMV},
description={Adaptive Predicted Mean Vote}
}

\newglossaryentry{pmvs}
{
name={PMV$_{SET}$},
description={PMV calculated using the SET temperature inplace of $t_{db}$}
}

\newglossaryentry{pmvg}
{
name={PMV$_{Gagge}$},
description={PMV calculated using the two-node heat balance model}
}

\newglossaryentry{db2}
{
name={Comfort DB},
description={ASHRAE Global Thermal Comfort Database II}
}

\makenoidxglossaries