### 2024-08-15 Manuscript/src/myglossary.tex
feat: code to analyse data from studies with three air speed measurements

\usepackage[nonumberlist,nogroupskip]{glossaries}
% \usepackage{glossaries}

\setglossarystyle{super}

\newglossaryentry{7730}
{
name={ISO~7730:2005},
description={ISO~7730:2005 is a thermal comfort standard developed by ISO}
}

\newglossaryentry{55}
{
name={ASHRAE~55:2023},
description={ASHRAE~55:2023 is a thermal comfort standard developed by ANSI and ASHRAE}
}

% \makenoidxglossaries