### 2024-12-05 Manuscript/src/main.tex
submission: first submission energy and buildings

%! Author = sbbfti
%! Date = 10/06/2020

\documentclass[review, 1p]{elsarticle}

\usepackage{mypreamble}
\usepackage{setspace}

\input{myglossary.tex}

\begin{document}

    \input{sections/titleAbstract}

    \input{myacronyms.tex}

    \doublespacing

    \linenumbers

    \acresetall

    \input{sections/introduction}

    \input{sections/methodology}

    \input{sections/results}

    \input{sections/conclusions}

    \clearpage

    \bibliography{mybibfile}

    \clearpage

    \input{sections/appendix}

\end{document}