### 2025-01-16 Manuscript/src/myglossary.tex
review: calculated Icl,r ISO, reviewed the paper until the start methodology.tex, changed V to Vr, Icl to Icl,r

\usepackage[nonumberlist,nogroupskip]{glossaries}

\setglossarystyle{super}

\newglossaryentry{7730}
{
name={ISO~7730:2005},
description={ISO~7730:2005 is a thermal comfort standard developed by ISO}
}

\newglossaryentry{55}
{
name={ASHRAE~55:2023},
description={ASHRAE~55:2023 is a thermal comfort standard developed by ANSI and ASHRAE}
}

\newglossaryentry{9920}
{
name={ISO~9920:2007},
description={ISO~9920:2007 is a thermal comfort standard developed by ISO}
}